Practice problem set 2

Problem 3

Simulate 200 values from $X_t = \sin(\frac(t)(20) + Z_t$ where $Z_t$ iid
with mean 0 and variance 0.25.

Might be interesting to use different distributions for x.

```{r}
n = 200

t = 1:n
z = rnorm(n, sd = sqrt(0.25))
x = sin(t / 20) + z 

plot(t, x)
```

Check for whiteness using sample ACF.

```{r}
acf(x)

# Verify location of the confidence interval line
2 / sqrt(n)
```

All of these are over the noise variance, so we conclude that this is not
white noise.

The Portmanteau test:

```{r}
Box.test(x, lag=20, type='Ljung')
Box.test(x, lag=3, type='Ljung')
```
TODO- what's the significance of the lag?
How to pick the right one?

Tiny p-value.

## Problem 5

```{r}
n = 200
x = arima.sim(list(ma=0.5), n=n)
mean(x)
plot(x)

# Sample ACF is right on- 2nd entry should be 0.4
acf(x)

# Theoretical 95% confidence interval for the mean
alpha = 0.05
interval = qnorm(1 - alpha / 2) * sqrt((9 * n - 4) / (4 * n^2))

# Simulate a bunch of sample means
sims = replicate(1e4, mean(arima.sim(list(ma=0.5), n=n)))
hist(sims)

# How many were inside the theoretical confidence interval?
sum(abs(sims) < interval) / length(sims)
```

Bumping it up to 100K replications gives 0.94932 inside the 95% confidence
interval. Exactly what we like to see.

Let's look at the ACF's for AR and MA processes. Try to develop a little
intuition here. I expect that an AR(p) process only has the ACF different
for the first p values of h. Nope- totally wrong!

```{r}
n = 200
x = arima.sim(list(ar=0.999), n=n)
plot(x)
acf(x)

x2 = arima.sim(list(ar=0.7), n=n)
acf(x2)
```

So this is in some sense 'close' to a random walk.

It looks like the MA(q) process should in fact have ACF different for the 
first q values of h. Yep.

```{r}
n = 200
x = arima.sim(list(ma=c(0.5, 2, 1)), n=n)
plot(x)
acf(x)
```

This is pretty cool.
