```{r}
library(xts)
library(MASS)
```

```{r}
source('functions.R')
load('fastrak.Rda')
fastrak = preclean(fastrak)
```

Focusing on a particular time series now.

Problem- the large model was not fully removing the seasonality and trend.
We'll address this by fitting an individual model to this station.

```{r}
santarosa = getstation(4300, fastrak)
```

```{r}
srfit = lm(count^(1/3) ~ year + month + weekday*hour, santarosa)
```

This is quite fast. I'm pleased.

```{r}
s = summary(srfit)
s$adj.r.squared
a = anova(srfit)
a
a['hour', 'Sum Sq'] / sum(a[, 'Sum Sq'])
```

This simple model explains the data well.
Almost all of the variation is due to the main effect of hour.
But the residuals have some serious problems.

```{r}
plot.lm(srfit, which=2)
```

```{r}
hist(residuals(srfit))
```

We suspect that this is due to sensor error. Let's trim the bottom 1% and
refit the model.

```{r}
cutoff = quantile(residuals(srfit), 0.01)
sr2 = santarosa[residuals(srfit) > cutoff, ]
dim(sr2)
```

```{r}
fit2 = lm(count^(1/3) ~ year + month + weekday*hour, sr2)
sr2$residuals = residuals(fit2)
```

```{r}
a2 = anova(fit2)
s2 = summary(fit2)
s2$adj.r.squared
```

# TODO
Need help interpreting this plot- is there any way to improve?
```{r}
plot.lm(fit2, which=2)
```

Still god-awful looking residuals on the qqplot. 
It's mostly symmetric now, but the tails are too heavy.
So we can't make inferences on this.

```{r}
hist(residuals(fit2))
```

Lets try the box cox method to find the proper transformation.

```{r}
fit3 = lm(count ~ year + month + weekday*hour, sr2)

boxcox(fit3)
```

This points directly to the cube root transform that we've been using.

```{r}
diffsr = diff(sr2$time)
islong = longrun(diffsr, 1, 1000)

santarosa = sr2[islong, ]
dim(santarosa)
#save(santarosa, file='santarosa.Rda')
```

Kind of interesting- we now have 6 periods where there are between 1000 and
1500 consecutive quality observations. An interesting analysis would be to
look at the ACF / PACF plots of each.

Make sure it's looking reasonable.


```{r}
diffsr = diff(sr2$time)
islong = longrun(diffsr, 1, 1400)

santarosa = sr2[islong, ]
dim(santarosa)
```

```{r}
with(santarosa[1000:1479, ], plot(time, residuals, type='l'))
```

These don't look too bad at all. What kind of ACF / PACF can we see?

```{r}
acf(santarosa$residuals)
pacf(santarosa$residuals)
```

Clearly point to an AR(1), and possibly with an AR(24) term.
Let's try some AIC model selection to fit.

```{r}
res = santarosa$residuals
plot(res, type='l')
mean(res)
```

We see some spikes where the model way underpredicted. Maybe this was a
special event.

```{r}
ar1 = ar(res, order.max=24)
ar1$aic
ar2 = ar(res)
ar2$aic
```

The default model selection is choosing 31 terms in the AR model.
Interesting. What about an ARMA model?

```{r}
arma1 = arima(res, order=c(2, 0, 2))
arma1$aic
arima(res, order=c(1, 0, 1))$aic
arima(res, order=c(1, 0, 2))$aic
arima(res, order=c(2, 0, 1))$aic
arima(res, order=c(2, 0, 2))$aic
arima(res, order=c(3, 0, 1))$aic 
arima(res, order=c(2, 0, 3))$aic # Best
arima(res, order=c(1, 0, 3))$aic
arima(res, order=c(3, 0, 3))$aic

# Not converging:
arima(res, order=c(4, 0, 4))$aic
arima(res, order=c(4, 0, 3))$aic
arima(res, order=c(3, 0, 4))$aic
```

```{r}
arma = arima(res, order=c(2, 0, 3))
arma$coef
```

