
Used the department servers to fit a mixed effects model. I'm hoping that
removed the trend and seasonality.

The model was:

```
fit2 = lmer(sqrt(count) ~ hour + weekday + month + year + (1|station) +(1|station:weekday), fastrak)
```

```{r}
load('fastrak.Rda')
```

Lets focus on a single time series now. We'll look at the one on Carquinez
bridge because that's the one I drive on most. This is I-80 East headed
away from the Bay Area. 

This brings up another very interesting possibility for modeling- most
of these time series come in pairs- toll plazas on both sides of the
freeway. Would it be possible to capture this relation in a statistical
model? Since typically traffic flows more heavily one way in the morning
and so on.

Look this information up from the station table.

```{r}
t1 = fastrak[fastrak$station == 1100, ]
# Sort on time
t1 = t1[order(t1$time), ]
```

Estimate total revenue over the period in billions of dollars.

```{r}
sum(fastrak$count) * 5 / 1e9
```

How many zeros are in the data?

```{r}
sum(t1$count == 0) / nrow(t1)
```

0.4?! No way. So we have significant missing data here which has been
encoded with zeros. How about in the whole data set?

```{r}
sum(fastrak$count == 0) / nrow(fastrak)
```

About a quarter. So the hard thing now is that some of these are in fact
true zeros. How can we distinguish? We could see if there's a gap between
the zeros and the next common observations.

```{r}
with(fastrak, hist(count[count < 100]))
h = hist(fastrak$count, breaks=c(-1, 1, 100, 1000, 1e4))
plot(h$counts)
```

A first look at a slice:

```{r}
slice = 25000:25100
t1s = t1[slice, ]
plot(t1s$time, t1s$count, type='l')
plot(t1s$time, t1s$residuals, type='l')
```

From this plot it seems like the zeros occur together for many times.
A simple way to detect this would be to count the number of consecutive
zeros- if it's greater than some number k then we encode all the values to
NA within that period.

```{r}
t1z = t1$count == 0
r1 = rle(t1z)
plot(r1$lengths[r1$lengths < 24])
```

We observe many cases of runs of zeros just being equal to 1 or 2. 
Those may be legitimate. Imagine a 5 hour period on a major freeway when no
cars went by. This seems highly improbable. Let's look at a 6 hour stretch
and see what makes sense.

```{r}
t1s = t1[50:150, ]
plot(t1s$time, t1s$count, type='l')
```

Here we see some 0 periods occuring at mid day on Tuesday and Wednesday- so
they're almost certainly invalid readings.
This motivates setting k = 3 or 4.

Here's a slice without zeros:

```{r}
slice = 15000:15100
t1s = t1[slice, ]
plot(t1s$time, t1s$count, type='l')
```

Let's write a function to filter these outliers.

```{r}
zerorun = function(x, k){
    # Determine whether the zeros in x occur in a run of 
    # at least k.
    # Returns a logical vector which is TRUE if the corresponding 
    # entry in x is in such a run
    iszero = x == 0
    r1 = rle(t1z)
    islong = (r1$lengths >= k)
    rep(islong, r1$lengths)
}

x = t1$count
k = 100

original = rle(x == 0)
length(original$values)
new = rle(zerorun(x, k))

original$lengths
length(new$values)

(new$lengths[new$values == FALSE])
max(new$lengths[new$values == FALSE])

d = data.frame(new$values, new$lengths)
head(d, 20)
d[d$new.values == FALSE, ]
```

