
```{r}
load('fastrak.Rda')
```

Maybe we should limit the scope of this to a single pair of stations. Which
ones have the most nonzero entries?

```{r}
totals = data.frame(totlen = tapply(fastrak$count, fastrak$station, length)
, nozerolen = tapply(fastrak$count, fastrak$station, function(x) sum(x != 0)))

totals$ratios = with(totals, nozerolen / totlen)
totals$station = row.names(totals)

plot(totals$ratios)
```

We observe that one pair is perfect- 25000 and 25100. What are the
corresponding counts? Very small. Too bad.

```{r}
t2 = totals[order(totals$nozerolen, decreasing=TRUE), ]
t2
t2[1:30, ]
```

If we want to first just focus on a pair of stations it looks like 4300 and
4301 (Santa Rosa) would be a good bet. Also 15900 and 15901 (101).

How about doing the Golden Gate bridge? That's 5001 and 5100

```{r}
totals[totals$station %in% c(5001, 5100), ]
```
40% nonzero. That's terrible, and similar to the Carquinez bridge.

Bay bridge?
```{r}
totals[totals$station %in% c(17300), ]
```

90%. This is the first one that seems reasonable. We can check some plots
and see if we get the same sort of 0 spots.

