
```{r}
load('fastrak.Rda')
```

Maybe we should limit the scope of this to a single pair of stations. Which
ones have the most nonzero entries?

```{r}
totals = data.frame(totlen = tapply(fastrak$count, fastrak$station, length)
, nozerolen = tapply(fastrak$count, fastrak$station, function(x) sum(x != 0)))

totals$ratios = with(totals, nozerolen / totlen)
totals$station = row.names(totals)

plot(totals$ratios)
```

We observe that one pair is perfect- 25000 and 25100. What are the
corresponding counts? Very small. Too bad.

```{r}
t2 = totals[order(totals$nozerolen, decreasing=TRUE), ]
t2
t2[1:30, ]
```

If we want to first just focus on a pair of stations it looks like 4300 and
4301 (Santa Rosa) would be a good bet. Also 15900 and 15901 (101).

```{r}
sr = fastrak[fastrak$station == 4300, ]
with(sr[sr$year == 2010, ], plot(time, count))
```
Everything is showing this doubled period in July of 2010. What if we just
drop these suspicious observations? Also, it would be feasible to run
through and plot most stations, dropping suspicious areas as we see fit. This would let us
still run the ANOVA model.

They all seem to be missing a chunk in July 2010 as well as a zero period
in October of 2009.
```{r}
library(scales)
bigplot = function(stationnumber){
    # Plot the whole series for a station
    x = fastrak[fastrak$station == stationnumber, ]
    with(x, plot(time, count, col=alpha('black', 0.1)))
}
# This one is missing the July chunk
bigplot(21200)
# This one has more problems
bigplot(300)
# More problems
bigplot(2501)
# Missing july chunk
bigplot(6701)
# Missing some more chunks
bigplot(17500)
```

How about doing the Golden Gate bridge? That's 5001 and 5100

```{r}
bigplot(5100)
totals[totals$station %in% c(5001, 5100), ]
```
40% nonzero. That's terrible, and similar to the Carquinez bridge.

Bay bridge?
```{r}
bigplot(17300)
totals[totals$station %in% c(17300), ]
```

90%. This is the first one that seems reasonable. We can check some plots
and see if we get the same sort of 0 spots.

