```{r}
load('fastrak.Rda')
```

Ok, focusing on the Bay Bridge toll plaza now.

```{r}
bb = fastrak[fastrak$station == 17300, ]
bb = bb[order(bb$time), ]
dim(bb)
```

Check if we're seeing these runs of zeros.

```{r}
zeros = bb$count == 0
z = rle(zeros)
hist(z$lengths[z$lengths < 24])
sort(z$lengths, decreasing=T)
```

I find it hard to believe that there were ever any hours when the Bay
Bridge was open without traffic.

```{r}
with(bb, plot(time, count))
```

Now this looks totally ridiculous. There's lots of missing data and the
count is surely greater than this.

```{r}
max(bb$count)
```

```{r}
with(bb[1:2000, ], plot(time, count))

with(bb[1:5000, ], plot(time, count))
with(bb[5000:10000, ], plot(time, count, alpha=0.2))
with(bb[10000:15000, ], plot(time, count))
with(bb[15000:20000, ], plot(time, count))
```

I find it very hard to believe that these maxima are really around 500.
1 car a second would be 3600 / hour. 500 a second means the max rate per
day is around 1 car in every 14 seconds. That's too few.


```{r}
with(bb, plot(time, count))
```

```{r}
# plot in chunks of 5000
plotpart = function(k, dframe=bb, chunk=5000){
    with(dframe[k*chunk + (1:chunk), ], plot(time, count))
}
plotpart(4)
plotpart(5)
plotpart(6)
plotpart(7)
plotpart(8)
plotpart(9)
plotpart(10)
```

Is it all this bad at the other stations? Say San Mateo?

```{r}
sm = fastrak[fastrak$station == 15900, ]
sm = sm[order(sm$time), ]
dim(sm)
```

```{r}
for (i in 1:11){ 
    plotpart(i, dframe=sm) 
    Sys.sleep(5)
}
```
This one looks quite a bit more consistent than the Bay bridge.

```{r}
with(sm, plot(time, count))

with(sm[sm$year == 2010, ], plot(time, count))
```

So maybe we just pick a consistent period out of here. I wonder if the
other one looks the same?

```{r}
sm2 = fastrak[fastrak$station == 15901, ]
sm2 = sm2[order(sm2$time), ]
dim(sm2)
```

```{r}
with(sm2, plot(time, count))
```

```{r}
plotpart(80, sm2, chunk=500)
```

What about just throwing out the zeros?

```{r}
with(sm, hist(count[count < 10]))
```

I'd say this provides pretty strong evidence for coding 0 as NA.

```{r}
zsm = rle(sm$count == 0)
max(zsm$lengths)
hist(zsm$lengths[zsm$lengths < 10])
```

Let's look at aggregated counts:

```{r}
# This is an expensive operation
totals = tapply(fastrak$count, fastrak$time, sum)
tot = data.frame(total = totals, time = as.POSIXct(names(totals)))
sapply(tot, class)
```

Plot it with some nice alpha transparency and a linear model.

@final
```{r}
library(scales)
tfit = lm(total ~ time, tot)
png('totals.png')
with(tot, plot(time, total, col=alpha('black', 0.1)))
abline(tfit, col='blue', lwd=2)
dev.off()
```

A linear model.

```{r}
summary(tfit)
```


